### 2018-03-20 CO749-GraphColoring/assignments/A4.tex
solved it

	Let $G$ be a plane graph, let $S$ be a connected subgraph of $G$, and let $L$ be a $5$-list-assignment for $G$. If $G$ is $S$-critical with respect to $L$ then
	$$v(G|S) \leq c(\defi(G|S) - 1),$$
	where $\defi(G|S) = e(G|S) - 3v(G|S).$
\end{theorem}
We will prove the following lemma, under the assumption that a minimum counterexample $G$ has a vertex in $V(G)\backslash V(S)$ with at least three neighbours in $S$.
\begin{lemma}
	Let $G$ be a plane graph with connected subgraph $S$ and let $L$ be a $5$-list-assignment for $G$. If $\phi$ is an $L$-coloring of $S$ that extends to an $L$-coloring of $G$ then
	$$\log_2 E(\phi) \geq (v(G|S) - (c+1)\defi(G|S))/9,$$
	where $E(\phi)$ is the number of extensions of $\phi$ to $G$ and $c$ is the constant of Theorem \ref{th:def}.
\end{lemma}
\begin{proof}
	Suppose not. Let $(G,S,L)$ be a minimum counterexample, minimizing $v(G|S)$. We may assume $G$ has a vertex $v$ in $V(G)\backslash V(S)$ with at least three neighbours in $S$.
	\paragraph{}

>>>>

	Let $G$ be a plane graph, let $S$ be a connected subgraph of $G$, and let $L$ be a $5$-list-assignment for $G$. If $G$ is $S$-critical with respect to $L$ then
	$$v(G|S) \leq c(\defi(G|S) - 1),$$
	where $\defi(G|S) = e(G|S) - 3v(G|S).$
We will prove the following lemma, under the assumption that a minimum counterexample $G$ has a vertex in $V(G)\backslash V(S)$ with at least three neighbours in $S$.
	Let $G$ be a plane graph with connected subgraph $S$ and let $L$ be a $5$-list-assignment for $G$. If $\phi$ is an $L$-coloring of $S$ that extends to an $L$-coloring of $G$ then
	$$\log_2 E(\phi) \geq (v(G|S) - (c+1)\defi(G|S))/9,$$
	where $E(\phi)$ is the number of extensions of $\phi$ to $G$ and $c$ is the constant of Theorem \ref{th:def}.
	Suppose not. Let $(G,S,L, \phi)$ be a minimum counterexample, minimizing $v(G|S)$. We may assume $G$ has a vertex $v$ in $V(G)\backslash V(S)$ with at least three neighbours in $S$. Let $H = S+v$. 
	We first consider the case that $v(G|S) = 1$, i.e. $G = S + v$.
	\frac{v(G|S) - (c+1)\defi(G|S)}{9} &= \frac{1 - (c+1)e(G|S) +3(c+1)}{9} \\
	&\begin{cases}
		=\frac{1}{9}, &\text{if $|N(v)\cap v(S)| = 3$}  \\
		\leq\frac{1-(c+1)}{9} <0, &\text{if $|N(v) \cap V(S)| \geq 4$}	\end{cases}
	Since $\phi$ has at least one extension to $G$ if $|N(v) \cap V(S)| \geq 4$ and $\log_2(1) = 0$, and $\phi$ has at least two extensions to $G$ if $|N(v) \cap V(S)| = 3$ (as at least two colors in $v$'s list are not used at neighbours) and $\log_2(2) = 1 > \frac{1}{9}$, the desired bound holds.
	Therefore we may assume that $v(G|S) \geq 2$. Consider the case that $|N(v) \cap V(S)| \geq 4$. Let $\phi'$ be an extension of $\phi$ to $G$, restricted to $H$. Then $\phi'$ is an $L$-coloring of $H$ which extends to $G$. By minimality, 
	$$\log_2 E(\phi') \geq \frac{v(G|H) - (c+1)\defi(G|H)}{9}$$
	Now $v(G|H) = v(G|S) - 1$. So we see that
	\frac{v(G|H) - (c+1)\defi(G|H)}{9} &= \frac{v(G|S) - 1 -(c+1)\defi(G|S) +(c+1)|N(v)\cap V(S)| - 3(c+1)}{9} \\
	&\geq \frac{v(G|S) - 1 -(c+1)\defi(G|S) +(c+1)4 - 3(c+1)}{9}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S)}{9}.
	Hence since $E(\phi) \geq E(\phi')$ this yields the desired bound.
	Thus we may assume that $|N(v) \cap V(S)| = 3$. Yet again we distinguish cases. First suppose there exists $\phi_1$ and $\phi_2$ distinct $L$-colorings of $H$, extended from $\phi$, which extend to $G$. Without loss of generality suppose $E(\phi_1) \leq E(\phi_2)$. Then, using minimality,
	\log_2E(\phi) &\geq \log_2(E(\phi_1) + E(\phi_2)) \\
	&\geq \log_2(2E(\phi_1))\\
	&= 1 + \log_2(E(\phi_1))\\
	&\geq 1+ \frac{v(G|H) - (c+1)\defi(G|H)}{9}\\
	&= 1 + \frac{v(G|S) - 1 -(c+1)\defi(G|S) +(c+1)|N(v)\cap V(S)| - 3(c+1)}{9}\\
	&= 1 + \frac{v(G|S) -(c+1)\defi(G|S)}{9} + \frac{-1}{9} \\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S)}{9}.
	Thus yielding the desired bound.
	Therefore, finally, we suppose without loss of generality $\phi_1$ and $\phi_2$ are distinct $L$-colorings of $H$ and $\phi_1$ extends to $G$, but $\phi_2$ does not. This is last case to consider as at least one $L$-coloring of $H$ must extend to $G$ as $\phi$ does.
	Let $G' \subseteq G$ be a subgraph of $G$ that is $H$-critical with respect to $\phi_2$. Such $G'$ exists since $\phi_2$ does not extend to $G$. By Theorem \ref{th:def} $$v(G'|H) \leq c(\defi(G'|H) -1).$$
	Let $\phi'$ now denote the restriction of an extension of $\phi$ to $G$ to $G'$. That is, $\phi'$ is an $L$-coloring of $G'$ that extends to $G$. So by minimality
	$$\log_2(E(\phi')) \geq \frac{v(G|G') - (c+1)\defi(G|G')}{9}.$$
	It is easy to see that $v(\cdot)$ and $\defi(\cdot)$ satisfy the ``chain rule", that is
	$$\defi(G|H) = \defi(G|G') + \defi(G'|H)$$
	$$v(G|S) - 1 = v(G|H) = v(G|G') + v(G'|H).$$
	We make one final observation before the final bound
	$$\defi(G|H) = \defi(G|S) |N(v)\cap V(S)| -3 = \defi(G|S) $$
	Now we see that
	\log_2E(\phi) &\geq \log_2E(\phi')\\
	&\geq \frac{v(G|G') - (c+1)\defi(G|G')}{9} \\
	&= \frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|H) + (c+1)\defi(G'|H)}{9}\\
	&=\frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|S) + (c+1)\defi(G'|H)}{9}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +\defi(G'|H) + c - 1}{9} \\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +v(G'|H)/c + 1 + c - 1}{9}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S)}{9},
